---
title: "GRD 610A Data Visualization II"
subtitle: "Labels"
author: "Jenn Schilling"
date: "February 24, 2021"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, bottom-note.css]
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

options(knitr.duplicate.label = "allow")

knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)
 
```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 80%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 5: Basic Principles of Data Visualization of *The Truthful Art* (Cairo)  

- Lab on basics of mapping and options for visualizing data on a map - Chapter 7 of *Data Visualization* (Healy)  

- Homework Assignment #4


---

# Of Conjectures and Uncertainty

### Conjecture

**Definition:** "a guess about something based on how it seems and not on proof" (Cambridge Dictionary)

**Requirements** (Cairo, 2016, pp. 102-104)
- Makes sense
- Testable
- Made of several components
- If any component changes, the conjecture becomes useless

???

"Conjectures first need to make sense (even if they end up being wrong) based on existing knowledge of how nature works." (Cairo, 2016, p. 102)  

"For a conjecture to be good, it also needs to be testable...Being testable also implies being falsifiable." (Cairo, 2016, p. 103)  

"A good conjecture is made of several components, and these need to be hard to change without making the whole conjecture useless. In the words of physicist David Deutsch, a good conjecture is "hard to vary, because all its details play a functional role." (Cairo, 2016, p. 104)



---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "25%",
          warn_when = 60)
```

---

## Chapter 5: Graph Tables, Add Labels, Make Notes

.center[
```{r first-labels, echo = FALSE}

by_country <- organdata %>%
  group_by(consent_law, country) %>%
  summarize_if(is.numeric, 
               list(mean = mean, sd = sd), # note funs is deprecated
               na.rm = TRUE) %>% 
  ungroup()

p <- ggplot(data = by_country,
            mapping = aes(x = roads_mean,
                          y = donors_mean)) 

p + 
  geom_point() + # plot points
  geom_text(mapping = aes(label = country)) # plot the labels

```
]


---


```{r annotate, include = FALSE}
p <- ggplot(data = organdata,
            mapping = aes(x = roads,
                          y = donors))

p + 
  geom_point() +
  annotate(geom = "rect", 
           xmin = 125, xmax = 155, 
           ymin = 30, ymax = 35, 
           fill = "red", alpha = 0.2) + 
    annotate(geom = "text", 
           x = 91, 
           y = 33, 
           label = "A surprisingly high \n recovery rate", 
           hjust = 0) 

```


`r chunk_reveal(chunk_name = "annotate", title = "#Annotations", widths = c(50,50))`

---

class: center, middle, inverse

# Lab Time

### Chapter 7



---

# Homework Assignment

**Task: ** Create 3 maps  

**Due: ** March 3, 2021 

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42755_1&course_id=_35307_1)

**Notes**

- NOTES ABOUT DATA TO USE

- Use your resources: Healy, Google, Student Community BUT cite where you get code from if you copy it directly  

- See Blackboard assignment for a template `.Rmd` file

---

# Tasks to Complete

- Reading (Cairo - Chapter 10: Mapping Data)

- Homework #4




