---
title: "GRD 610A Data Visualization II"
subtitle: "Final Project 3"
author: "Jenn Schilling"
date: "April 21, 2021"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, bottom-note.css]
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

options(knitr.duplicate.label = "allow")

knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)
library(ggrepel)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Montserrat"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 70%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 11: Uncertainty and significance in *The Truthful Art* (Cairo)  

- Final Project



---
# Recall From Prior Chapters

- Hypothesis

- Mean

- Standard deviation

- Standardized scores (z-scores)

- Normal Distribution

---
# Terms

- **Standard Error** - measures how far the mean of the sample might be from the mean of the population; calculated by dividing the standard deviation by the square root of the sample size 

- **Confidence Interval** - the uncertainty around a statistic, based on a confidence level (i.e. 95%); calculated using the point estimate, critical value, and standard error

- **Critical Value** - a number based on the confidence level and either the normal distribution or a t-distribution, depending on sample size 


???

Sample size - (number of samples - 1)

Confidence interval - we are 95% confident that the values fall within this range i.e. 95 out of 100 times we will get a mean within this range

Figure 11.6 for confidence intervals and paragraph at bottom of p. 309 for critical value

Figure 11.7 for t-values

Figure 11.9 shows how to calculate confidence intervals for percentages

---
# Disclosing Uncertainty

- How can you convey uncertainty in a data visualization?

???

Ideas: 
put confidence interval in notes
use error bars on a dot plot, bar plot, or line (Figurer 11.10)
use a gradient plot (Figure 11.12)
use a violin plot (Figure 11.11)
use a fanplot (R package) (Figure 11.13)
use error area on a map (Figure 11.14)

---
# Questions to Ask

"the secret behind any successful data project is asking the people who know a lot about the data at hand and its shortcomings, about how it was garnered, processed, and tested."

"If you have downloaded data from public sources...**don't publish anything: take your hunches to experts**"

Ask: How has stuff been measured?

From Heather Krause:
- What is being measured?
- Who is the population being measured?
- What time period is being measured?

.bottom-note[
(Cairo, 2016, p. 316, 319)
]

???

- Read the documentation about the data and make sure you understand how the data has been acquired, what is being measured, who the data covers, and over what time period

---
# Some More Terms

- **p-value** - the probability of obtaining the data or more extreme given that the null hypothesis is true

- **Null Hypothesis** - there is no effect or difference

- **Experimental Group** - subset of the sample who receive the treatment

- **Control Group** - subset of the sample who do not receive the treatment (stays the same)

- **Statistically Significant** - the result is unlikely to occur just by chance

- **Effect Size** - measures the strength of the relationship between two variables

- **Statistical Power** - the probability of observing an effect of a certain size 

**Key takeaway:** When trying to understand or draw a conclusion based on data, be sure to look at more than just the significance. 

---
# Discussion

- What is your main takeaway from this chapter?

--

- Which figures stood out to you? Why?

--

- How will you use the information in this chapter when creating data visualizations?

--

- What questions do you have?


---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "25%",
          warn_when = 60) 
```

---

# Final Project

**Task: ** Using R and an approved dataset, create a polished infographic that contains at least 5 different charts.  

**Due: ** May 5, 2021

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42757_1&course_id=_35307_1)

- Pick a dataset using one of the resources available on Blackboard. The dataset must be approved; email Jenn a link to the dataset for approval prior to **March 31**.  

- Create summary statistics and 5-7 polished visuals based on the dataset using **only R**. 

- The visuals created in R must include: titles, axis labels, data labels and legends (as needed), your name and the data source (as footnotes/captions)  

- Output the visuals to `.svg` and import them into Adobe Illustrator to create an infographic that includes a narrative, the visuals from R, and other graphics.  

- Turn in your R code as an `.Rmd` file (including comments about what you did and why), your data file, and the final infographic. 

- In class on **April 28**, be prepared to share information about your dataset, your visuals, what they show, and the choices you made in their design. You will share both your code and the final infographic.

???

In class on April 28, we will be sharing critiques of each others work. 

Presentation to the class: share your .Rmd file and be prepared to explain your data (what is it about, what does it tell you, what messages did you find in it), your visualizations (how did you create them, why did you create them, what design choices you made); share your infographic (what design choices did you make and why, what story or message does the infographic show)

---

# Tasks to Complete

- Reading (Cairo - Chapter 12: On Creativity and Innovation)

- Work on Final Project



