---
title: "GRD 610A Data Visualization II"
subtitle: "Final Project 3"
author: "Jenn Schilling"
date: "April 21, 2021"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, bottom-note.css]
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

options(knitr.duplicate.label = "allow")

knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)
library(ggrepel)


```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Montserrat"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 70%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 11: Uncertainty and significance in *The Truthful Art* (Cairo)  

- Final Project



---
# Recall From Prior Chapters

- Hypothesis
- Mean
- Standard deviation
- Standaradized scores (z-scores)
- Normal Distribution

---
# Terms

- **Standard Error** 


---
# Discussion

- What is your main takeaway from this chapter?

--

- Which figures stood out to you? Why?

--

- How will you use the information in this chapter when creating data visualizations?

--

- What questions do you have?


---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "25%",
          warn_when = 60)
```

---

# Final Project

**Task: ** Using R and an approved dataset, create a polished infographic that contains at least 5 different charts.  

**Due: ** May 5, 2021

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42757_1&course_id=_35307_1)

- Pick a dataset using one of the resources available on Blackboard. The dataset must be approved; email Jenn a link to the dataset for approval prior to **March 31**.  

- Create summary statistics and 5-7 polished visuals based on the dataset using **only R**. 

- The visuals created in R must include: titles, axis labels, data labels and legends (as needed), your name and the data source (as footnotes/captions)  

- Output the visuals to `.svg` and import them into Adobe Illustrator to create an infographic that includes a narrative, the visuals from R, and other graphics.  

- Turn in your R code as an `.Rmd` file (including comments about what you did and why), your data file, and the final infographic. 

- In class on **April 28**, be prepared to share information about your dataset, your visuals, what they show, and the choices you made in their design. You will share both your code and the final infographic.

???

In class on April 28, we will be sharing critiques of each others work. 

Presentation to the class: share your .Rmd file and be prepared to explain your data (what is it about, what does it tell you, what messages did you find in it), your visualizations (how did you create them, why did you create them, what design choices you made); share your infographic (what design choices did you make and why, what story or message does the infographic show)

---

# Tasks to Complete

- Reading (Cairo - Chapter 12: On creativity and innovation)

- Work on Final Project



