---
title: "GRD 610A Data Visualization II"
subtitle: "Case Studies"
author: "Jenn Schilling"
date: "March 31, 2021"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, bottom-note.css]
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

options(knitr.duplicate.label = "allow")

knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)
library(ggrepel)

```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Montserrat"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 70%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 7: Visualizing Distributions in *The Truthful Art* (Cairo)  

- Lab on the analysis and building of a few visualizations; additional plot types. 

- Homework Assignment #7



---
# Visualizing Distributions

**Why visualize distributions?**
--

- A central measure does not usually represent the data

- A distribution gives the viewer a better perspective and understanding of the underlying data.

--

- *"Lesson learned, again: Never trust just one statistic. Always plot your data."* (Cairo, 2016, p. 175)

???

The previous chapter covered measures of central tendency; this chapter covers measures of spread

Distributions can be visualized with box plots, histograms, and density plots; as we learned during the lab in Week 9

---
# Terms

**Variance** - a statistic that measures the spread of a dataset; a statistic that measures the distance between a set of numbers and their mean value  


**Standard Deviation** - a statistic that measures the spread of a dataset; the square root of the variance; has the same units as the data  


**z-score** - a measure of the number of standard deviations between a data point and the mean  


**Standardization** - putting data on the same scale to compare it; calculating z-scores or rates per 100,000 are examples of standardization


---
# Visualizing Distribution/Frequency

- Histograms (Figure 7.11)  
- Bar graphs (Figure 7.12)  
- Strip plot (Figures 7.10, 7.13)  
- Maps (Figure 7.13)  
- Network graphs (Figure 7.14)  

What other ways can you think of?

---
# An Important Note

**Small population => more variation** 

When there are fewer people/samples, there is more variation in the values and more values towards the extremes

???

When there is a larger population, there is less variation and the data is more reliable  

Be careful when calculating rates (or drawing conclusions) from a small population  

---
# Normal Distribution

**Properties**
- Mean, Median, and Mode are the same  
- Symmetrical (50% below the mean; 50% above the mean)  
- The percentage of scores between certain standard deviations from the mean are known (i.e. 95.4% are within 2 standard deviations)

**Why does this matter?**  
Many metrics follow approximately normal distributions, which makes it possible to calculate the probability of any data point, given the mean and standard deviation.

---
# Percentiles

- Another way to measure spread of a dataset

- The median is the 50th percentile of a dataset; it is in the middle of the distribution.  

- The median and other percentiles are **resistant** statistics, meaning they are not affected by outliers or extreme values (as the mean and standard deviation are).  

- **Quartiles** divide a dataset into quarters (25th, 50th, and 75th percentiles)  

- Visualize quartiles using **box-and-whisker plots**

- **Interquartile Range** - the distance/difference between the 25th and 75th percentiles (first and third quartiles); used to compute the whiskers on a box-and-whisker plot (each whisker extends 1.5 times the interquartile range below the first quartile or above the third quartile)

- Why use box-and-whisker plots?

???

Use box-and-whisker plots because:  
- shows less detail than a histogram/dot plot/etc. and are less likely to obscure important information  
- uses quartiles which means we can see portions of equal size (each is 25% of the data)  
- highlights outliers  
- makes it easier to compare distributions  

---
# Discussion

- Why examine measures of central tendency, spread, and shape?

--

- What stood out to you in this chapter?

--

- What can you apply from this chapter when creating data visualizations?

--

- What questions do you have?


---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "25%",
          warn_when = 60)
```



---

class: center, middle, inverse

# Lab Time

### Sections 8.0 - 8.4

---

# Homework Assignment

**Task: ** Create 3 charts using the new plot types from class 

**Due: ** April 7, 2021 

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42755_1&course_id=_35307_1)

**Notes**

- Use your own dataset (start looking for a dataset that you could use for the Final Project or go back to the dataset from the Midterm Project)

- Use your resources: Healy, Google, Student Community BUT cite where you get code from if you copy it directly  

- See Blackboard assignment for a template `.Rmd` file

---

# Tasks to Complete

- Reading (Cairo - Chapter 8: Revealing Change)

- Homework #7

- Prepare for your Data Visualization of the Week




