---
title: "GRD 610A Data Visualization II"
subtitle: "Homework #XX"
author: "Jenn Schilling"
date: "2021-02-10"
output:
 html_document:
  theme: "simplex" # Theme Gallery: https://www.datadreaming.org/post/r-markdown-theme-gallery/
---


```{r setup, include=FALSE}

## By default, show code for all chunks in the knitted document,
## as well as the output. To override for a particular chunk
## use echo = FALSE in its options.
knitr::opts_chunk$set(echo = TRUE)

## Set the default size of figures
knitr::opts_chunk$set(fig.width=8, fig.height=5)  

## Load the libraries we will be using
library(gapminder)
library(here)
library(socviz)
library(tidyverse)

```



******

# Getting Data 

If you want to use a different dataset package, run one of the install.package() line once in the Console
```{r install-package, include=TRUE, eval=FALSE}

install.packages("palmerpenguins")

install.packages("babynames")

```

Then add the library function call to the setup section above
```{r add-library, include=TRUE, eval=FALSE}

library(palmerpenguins)

library(babynames)

```

If you want to use your own dataset, add the code below to the setup section

```{r use-your-own-data, include=TRUE, eval=FALSE}

# From a URL
 url <- "https://cdn.rawgit.com/kjhealy/viz-organdata/master/organdonation.csv"
 organs <- read_csv(file = url)

# From your computer (save the file in the same folder as your Rmd file)
 data <- read_csv("file_name.csv")

```

**Remove this section after you implement it (everything between and including the ```******```)** 

******

# Homework Assignment

This is where you will put your R code and create your plots

```{r first-plot}



```

```{r second-plot}



```

```{r third-plot}



```

