---
title: "GRD 610A Data Visualization II"
subtitle: "Case Studies"
author: "Jenn Schilling"
date: "March 31, 2021"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, bottom-note.css]
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

options(knitr.duplicate.label = "allow")

knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)
library(ggrepel)

sales <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/sales.csv') %>% mutate(country = ifelse(country == "WW", "World", country))

```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Montserrat"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 70%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 7: Visualizing Distributions in *The Truthful Art* (Cairo)  

- Lab on the analysis and building of a few visualizations; additional plot types. 

- Homework Assignment #7



---
# Visualizing Distributions

**Why visualize distributions?**
--

- A central measure does not usually represent the data

- A distribution gives the viewer a better perspective and understanding of the underlying data.

--

- *"Lesson learned, again: Never trust just one statistic. Always plot your data."* (Cairo, 2016, p. 175)

???

The previous chapter covered measures of central tendency; this chapter covers measures of spread

Distributions can be visualized with box plots, histograms, and density plots; as we learned during the lab in Week 9

---
# Terms

**Variance** - a statistic that measures the spread of a dataset; a statistic that measures the distance between a set of numbers and their mean value  


**Standard Deviation** - a statistic that measures the spread of a dataset; the square root of the variance; has the same units as the data  


**z-score** - a measure of the number of standard deviations between a data point and the mean  


**Standardization** - putting data on the same scale to compare it; calculating z-scores or rates per 100,000 are examples of standardization


---
# Visualizing Distribution/Frequency

- Histograms (Figure 7.11)  
- Bar graphs (Figure 7.12)  
- Strip plot (Figures 7.10, 7.13)  
- Maps (Figure 7.13)  
- Network graphs (Figure 7.14)  

What other ways can you think of?

---
# An Important Note

**Small population => more variation** 

When there are fewer people/samples, there is more variation in the values and more values towards the extremes

???

When there is a larger population, there is less variation and the data is more reliable  

Be careful when calculating rates (or drawing conclusions) from a small population  

---
# Normal Distribution

**Properties**
- Mean, Median, and Mode are the same  
- Symmetrical (50% below the mean; 50% above the mean)  
- The percentage of scores between certain standard deviations from the mean are known (i.e. 95.4% are within 2 standard deviations)

**Why does this matter?**  
Many metrics follow approximately normal distributions, which makes it possible to calculate the probability of any data point, given the mean and standard deviation.

---
# Percentiles

- Another way to measure spread of a dataset

- The median is the 50th percentile of a dataset; it is in the middle of the distribution.  

- The median and other percentiles are **resistant** statistics, meaning they are not affected by outliers or extreme values (as the mean and standard deviation are).  

- **Quartiles** divide a dataset into quarters (25th, 50th, and 75th percentiles)  

- Visualize quartiles using **box-and-whisker plots**

- **Interquartile Range** - the distance/difference between the 25th and 75th percentiles (first and third quartiles); used to compute the whiskers on a box-and-whisker plot (each whisker extends 1.5 times the interquartile range below the first quartile or above the third quartile)

- Why use box-and-whisker plots?

???

Use box-and-whisker plots because:  
- shows less detail than a histogram/dot plot/etc. and are less likely to obscure important information  
- uses quartiles which means we can see portions of equal size (each is 25% of the data)  
- highlights outliers  
- makes it easier to compare distributions  

---
# Discussion

- Why examine measures of central tendency, spread, and shape?

--

- What stood out to you in this chapter?

--

- What can you apply from this chapter when creating data visualizations?

--

- What questions do you have?


---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "25%",
          warn_when = 60)
```

---
# Putting Plots Together

.pull-left[
**P1**

```{r plot-1, , fig.height = 5, fig.width = 5, echo = FALSE, eval = TRUE}

# Put data into long format to create tidy data
fredts_m <- fredts %>%
  select(date, sp500_i, monbase_i) %>% 
  gather(key = series,
         value = score,
         sp500_i:monbase_i)

# Line plot of the indexed data
p <- ggplot(data = fredts_m,
            mapping = aes(x = date, 
                         y = score,
                         group = series,
                         color = series))

p1 <- p + geom_line() +
  theme(legend.position = "top") +
  labs(x = "Date",
       y = "Index",
       color = "Series")

p1

```
]

.pull-right[
**P2**

```{r plot-2, fig.height = 5, fig.width = 5, echo = FALSE, eval = TRUE}

# Line plot of difference between the indexed series
# Note use of wide dataset to compute the difference
p <- ggplot(data = fredts,
            mapping = aes(x = date,
                          y = sp500_i - monbase_i))

p2 <- p + geom_line() +
  labs(x = "Date",
       y = "Difference")

p2

```
]

---
# Putting Plots Together `cowplot`

.pull-left[
```{r cowplot, eval = FALSE}

cowplot::plot_grid(p1, 
                   p2, 
                   nrow = 2, 
                   rel_heights = c(0.75, 0.25), 
                   align = "v")

```
]

.pull-right[
```{r output-cowplot, ref.label = 'cowplot', echo = FALSE, fig.height = 6, fig.width = 6}
```
]

.bottom-note[
[More about cowplot](https://wilkelab.org/cowplot/index.html)
]

---
# Putting Plots Together `patchwork`

.pull-left[
```{r patchwork, eval = FALSE}

library(patchwork)

p1 / p2 +
  plot_layout(nrow = 2,
              heights = c(0.75, 0.25))

```
]

.pull-right[
```{r output-patchwork, ref.label = 'patchwork', echo = FALSE, fig.height = 6, fig.width = 6}
```
]

.bottom-note[
[More about patchwork](https://patchwork.data-imaginist.com/)
]

---
# Move Axis Labels

.pull-left[
**Initial Plot**

```{r bar-plot-1, echo = FALSE, eval = TRUE, fig.width = 7, fig.height = 5}

xlabs <- c(0, 5, 10)

sales %>%
  filter(country == "World") %>%
  mutate(released = gsub("\\s*\\([^\\)]+\\)\\[[^\\)]+\\]", "", released)) %>%
  mutate(year = substr(released, nchar(released) - 3, nchar(released))) %>%
  ggplot(.) +
  geom_bar(aes(y = reorder(paste(title, paste0("(", year, ")"), sep = "\n"), 
                           sales), 
               x = sales,
               fill = artist), 
           stat = "identity") + 
  scale_fill_manual(values = c("#D4AF37", "red"),
                    guide = FALSE) +
  facet_wrap(~artist, scales = "free_y") +
  scale_x_continuous(labels = paste0("$", xlabs, "M"),
                     breaks = 1000000 * xlabs ) +
  labs(x = "",
       y = "",
       title = "Worldwide Album Sales") +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 12, hjust = 0),
        plot.title.position = "plot")

```
]

.pull-right[
`coord_cartesian(expand = FALSE)`

```{r bar-plot-2, echo = FALSE, eval = TRUE, fig.width = 7, fig.height = 5}

sales %>%
  filter(country == "World") %>%
  mutate(released = gsub("\\s*\\([^\\)]+\\)\\[[^\\)]+\\]", "", released)) %>%
  mutate(year = substr(released, nchar(released) - 3, nchar(released))) %>%
  ggplot(.) +
  geom_bar(aes(y = reorder(paste(title, paste0("(", year, ")"), sep = "\n"), 
                           sales), 
               x = sales,
               fill = artist), 
           stat = "identity") + 
  scale_fill_manual(values = c("#D4AF37", "red"),
                    guide = FALSE) +
  facet_wrap(~artist, scales = "free_y") +
  scale_x_continuous(labels = paste0("$", xlabs, "M"),
                     breaks = 1000000 * xlabs ) +
  labs(x = "",
       y = "",
       title = "Worldwide Album Sales") +
  theme_bw() +
  coord_cartesian(expand = FALSE) +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 12, hjust = 0),
        plot.title.position = "plot")


```
]

---
# Move Axis Labels

.pull-left[
**Initial Plot**

```{r show-bar-plot-1, ref.label = 'bar-plot-1', echo = FALSE, eval = TRUE, fig.width = 7, fig.height = 5}


```
]

.pull-right[
`scale_x_continuous(expand = c(0, 0))`

```{r bar-plot-2a, echo = FALSE, eval = TRUE, fig.width = 7, fig.height = 5}

sales %>%
  filter(country == "World") %>%
  mutate(released = gsub("\\s*\\([^\\)]+\\)\\[[^\\)]+\\]", "", released)) %>%
  mutate(year = substr(released, nchar(released) - 3, nchar(released))) %>%
  ggplot(.) +
  geom_bar(aes(y = reorder(paste(title, paste0("(", year, ")"), sep = "\n"), 
                           sales), 
               x = sales,
               fill = artist), 
           stat = "identity") + 
  scale_fill_manual(values = c("#D4AF37", "red"),
                    guide = FALSE) +
  facet_wrap(~artist, scales = "free_y") +
  scale_x_continuous(labels = paste0("$", xlabs, "M"),
                     breaks = 1000000 * xlabs,
                     expand = c(0, 0)) +
  labs(x = "",
       y = "",
       title = "Worldwide Album Sales") +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 12, hjust = 0),
        plot.title.position = "plot")


```
]

---

```{r data-prep, include = FALSE}
sales_dat <- sales %>%
  filter(country == "World") %>%
  mutate(released = gsub("\\s*\\([^\\)]+\\)\\[[^\\)]+\\]", "", released)) %>%
  mutate(year = substr(released, nchar(released) - 3, nchar(released))) %>%
  mutate(title_label = reorder(paste(title, paste0("(", year, ")"), sep = "\n"), 
                           sales))
```


# Gradient For Discrete Values

.pull-left[
```{r gradient-discrete, eval = FALSE}
  ggplot(data = subset(sales_dat, 
                       artist == "Taylor Swift"),
         mapping = aes(y = title_label,
                       x = sales,
                       fill = title_label)) +
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "RdPu", #<<
                    guide = FALSE) +
  facet_wrap(~artist, scales = "free_y") +
  scale_x_continuous(labels = paste0("$", xlabs, "M"),
                     breaks = 1000000 * xlabs,
                     expand = c(0, 0)) +
  labs(x = "",
       y = "",
       title = "Worldwide Album Sales") +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 12, hjust = 0),
        plot.title.position = "plot")
```
]

.pull-right[
```{r output-graident-discrete, ref.label = 'gradient-discrete', echo = FALSE, fig.height = 6, fig.width = 6}
```
]

.bottom-note[
[More about Color Brewer](https://colorbrewer2.org/)
]

---
# Manually Set Factor Order

.pull-left[
```{r factor-set}

sales_dat_ts <- sales_dat %>%
  filter(artist == "Taylor Swift") %>%
  mutate(title_label = 
           factor(title_label,
                  levels = c("Fearless\n(2008)",
                             "Speak Now\n(2010)",
                             "Red\n(2012)",
                             "1989\n(2014)",
                             "Reputation\n(2017)",
                             "Lover\n(2019)")))
 
```
]

.pull-right[
```{r plot-factor-set, echo = FALSE, fig.height = 6, fig.width = 6}
  ggplot(data = sales_dat_ts,
         mapping = aes(y = title_label,
                       x = sales,
                       fill = title_label)) +
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "RdPu", #<<
                    guide = FALSE) +
  facet_wrap(~artist, scales = "free_y") +
  scale_x_continuous(labels = paste0("$", xlabs, "M"),
                     breaks = 1000000 * xlabs,
                     expand = c(0, 0)) +
  labs(x = "",
       y = "",
       title = "Worldwide Album Sales") +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(size = 12, hjust = 0),
        plot.title.position = "plot")

```
]

---

class: center, middle, inverse

# Lab Time

### Section 8.5

### [R Graph Gallery](https://www.r-graph-gallery.com/)

---

# Homework Assignment

**Task: ** Create 3 charts using new plot types

**Due: ** April 7, 2021 

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42755_1&course_id=_35307_1)

**Notes**

- Use your own dataset (Final Project dataset)

- Use your resources: Healy, Google, Student Community BUT cite where you get code from if you copy it directly  

- See Blackboard assignment for a template `.Rmd` file

---

# Tasks to Complete

- Reading (Cairo - Chapter 8: Revealing Change)

- Homework #7

- Prepare for your Data Visualization of the Week




