---
title: "GRD 610A Data Visualization II"
subtitle: "Plot Types"
author: "Jenn Schilling"
date: "February 10, 2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
 
```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

# style_mono_light(base_color = "#23395b")

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 2 of *The Truthful Art* (Cairo)  

- Lab on Plot Types (Show the Right Numbers) - Chapter 4 of *Data Visualization* (Healy)  

- First Homework Assignment

---

# Discussion

The Five Qualities of Great Visualizations:
1. Truthful  
2. Functional  
3. Beautiful  
4. Insightful  
5. Enlightening  

From p. 45 of Cairo, A. (2016). *The truthful art: Data, charts, and maps for communication.* New Riders.

[Define the Qualities](https://padlet.com/jschilling_ccs/zmlnjzhiexx3kf14)

---

# Discussion

The Five Qualities of Great Visualizations:
1. Truthful  
2. Functional  
3. Beautiful  
4. Insightful  
5. Enlightening  

From p. 45 of Cairo, A. (2016). *The truthful art: Data, charts, and maps for communication.* New Riders.

--

**Do you agree that these are the five qualities of great visualizations?**

--

**Which is most important? Why?**

--

**Which is least important? Why?**

--

**What would you add?**


--

**What aspects of the data visualizations and infographics presented in Chapter 2 of *The Truthful Art* stood out to you?**



---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "27%",
          warn_when = 60)
```


---
# A Reminder of The ``ggplot`` Process

1. Tell the ``ggplot()`` function what our data is. (``data = ... ``)

2. Tell ``ggplot()`` what relationships we want to see.For convenience we will put the results of the first two steps in an object called ``p``.  (``mapping = aes(...)``)

3. Tell ``ggplot`` how we want to see the relationships in our data.  (``geom_...``)

4. Layer on geoms as needed, by adding them to the ``p`` object one at a time.  

5. Use some additional functions to adjust scales, labels, tick marks, titles. We’ll learn more about some of these functions shortly.

From p. 60 of Healy, K. (2019). *Data visualization: A practical introduction*. Princeton ; Oxford: Princeton University Press.

---

# Chapter 4: Show the Right Numbers

**"Code almost never works properly the first time you write it.** (Healy, 2019, p. 73)

---

class: center, middle, inverse

# Lab Time


---

# Homework Assignment

**Task: ** Create 3 different types of charts.  

**Due: ** February 17, 2020 

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42755_1&course_id=_35307_1)

**Notes**

- You should explore a dataset other than the *gapminder* one we've been using in the labs (some ideas: *babynames*, *palmerpenguins*, a CSV file you found)  

- To use an R package dataset, remember to run *install.packages("package_name")* once in the console and add *library(package_name)* to the *setup* portion of the .Rmd file

- See Blackboard assignment for a template .Rmd file

---

# Tasks to Complete

- Reading (see Syllabus)  

- Homework

- Prepare for your Data Visualization of the Week  


