---
title: "GRD 610A Data Visualization II"
subtitle: "Labels"
author: "Jenn Schilling"
date: "February 24, 2021"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, bottom-note.css]
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

options(knitr.duplicate.label = "allow")

knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)
 
```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 80%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 4: Of Conjectures and Uncertainty of *The Truthful Art* (Cairo)  

- Lab on adding axis labels, titles, legends, labeling points, and annotating plots - Chapter 5 of *Data Visualization* (Healy)  

- Homework Assignment #3

- Midterm 

---

# Of Conjectures and Uncertainty

### Conjecture

**Definition:** "a guess about something based on how it seems and not on proof" (Cambridge Dictionary)

**Requirements** (Cairo, 2016, pp. 102-104)
- Makes sense
- Testable
- Made of several components
- If any component changes, the conjecture becomes useless

???

"Conjectures first need to make sense (even if they end up being wrong) based on existing knowledge of how nature works." (Cairo, 2016, p. 102)  

"For a conjecture to be good, it also needs to be testable...Being testable also implies being falsifiable." (Cairo, 2016, p. 103)  

"A good conjecture is made of several components, and these need to be hard to change without making the whole conjecture useless. In the words of physicist David Deutsch, a good conjecture is "hard to vary, because all its details play a functional role." (Cairo, 2016, p. 104)

---

# Of Conjectures and Uncertainty

### Hypothesis

**Definition:** "a conjecture that is formalized to be tested empirically" (Cairo, 2016, p. 105)

**Variables**
- Independent (predictor/explanatory)
- Dependent (outcome/response)

.pull-left[
**Variable Scales**
- Nominal or Categorical
- Ordinal
- Interval
- Ratio
]
.pull-right[
**Variable Types**
- Discrete
- Continuous
]

???

When developing a hypothesis, you need to identify the specific variables and how they will be measured

Nominal - sex, location, political party  
Ordinal - ranking, happiness level  
Interval - incremental measurements of the same size, no zero point, IQ score (value can be twice as large, but not twice as intelligent)  
Ratio - meaningful zero point; weight, height, speed

---

.center[
```{r var-types-1, echo = FALSE, out.width = "95%"}
knitr::include_graphics("https://pbs.twimg.com/media/EkoCPz4VkAA_nST?format=jpg&name=4096x4096")
```
]

.bottom-note[Artwork by @allison_horst]

---

.center[
```{r var-types-2, echo = FALSE, out.width = "75%"}
knitr::include_graphics("https://pbs.twimg.com/media/EkoDNU6U0AISh-u?format=jpg&name=4096x4096")
```
]

.bottom-note[Artwork by @allison_horst]


---

# Of Conjectures and Uncertainty

### Study

"Test it [the hypothesis] against reality" (Cairo, 2016, p. 107)

**Types of Studies**
- Observational Study
- Cross-Sectional Study
- Longitudinal Study

**Population versus Sample**

**Extraneous Variables**
- Confounding variable
- Lurking variable

A **controlled experiment** goes beyond an observational study and tries to account for confounding variables.

???

Cross-sectional study - point in time 

We get data in a study from a sample and make inferences about a population (random sampling is best)  

Extraneous variables are variables that could affect the outcome but are not measured. Confounding variables can be identified and controlled for. Lurking variables are unknown so they cannot be identified. 

Controlled experiments include a large number of people that are representative of the population, divides people into an experimental group and a control group (blindly), measure both groups and compare the results.

---

# Of Conjectures and Uncertainty

### Uncertainty

"Data always vary randomly because the object of our inquires, nature itself, is also random."  
(Cairo, 2016, p. 114)

**Sample Variation** is the variation or change in a statistic or measurement from sample to sample.

???

Example - weighing yourself each day and the weight fluctuates

Example - random sampling of 1,000 people in U.S. and asking their political affiliation; get slightly different results each time

We can visualize uncertainty using shading and error bars/lines 

---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "25%",
          warn_when = 60)
```

---

## Chapter 5: Graph Tables, Add Labels, Make Notes

.center[
```{r first-labels, echo = FALSE}

by_country <- organdata %>%
  group_by(consent_law, country) %>%
  summarize_if(is.numeric, 
               list(mean = mean, sd = sd), # note funs is deprecated
               na.rm = TRUE) %>% 
  ungroup()

p <- ggplot(data = by_country,
            mapping = aes(x = roads_mean,
                          y = donors_mean)) 

p + 
  geom_point() + # plot points
  geom_text(mapping = aes(label = country)) # plot the labels

```
]

---

# Building a Labeled Plot

```{r plot-data}

library(ggrepel)

elections_historic %>% select(2:7)

```

---

# Building a Labeled Plot

```{r plot-prep}

p_title <- "Presidential Elections: Popular & Electoral College Margins"
p_subtitle <- "1824-2016"
p_caption <- "Data for 2016 are provisional"

x_label <- "Winner's share of Popular Vote"
y_label <- "Winner's share of Electoral College Votes"


```


```{r build-plot, include = FALSE}

p <- ggplot(data = elections_historic,
            mapping = aes(x = popular_pct,
                          y = ec_pct,
                          label = winner_label))

p + 
  geom_hline(yintercept = 0.5,
             size = 1.4,
             color = "gray80") +
  geom_vline(xintercept = 0.5,
             size = 1.4,
             color = "gray80") +
  geom_point() +
  geom_text_repel(seed = 1234) +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::percent) +
  labs(x = x_label,
       y = y_label,
       title = p_title,
       subtitle = p_subtitle,
       caption = p_caption)

```


---

`r chunk_reveal(chunk_name = "build-plot", title = "#Building a Labeled Plot", widths = c(50,50))`

???

Note seed set in geom_text_repel

---

# Label Outliers

Only label points with mean GDP greater than 25,000 OR mean health less than 1,500 OR the country Belgium

```{r outliers, eval = FALSE}
p <- ggplot(data = by_country,
            mapping = aes(x = gdp_mean,
                          y = health_mean))

p +
  geom_point() +
  geom_text_repel(data = subset(by_country, #<<
                                gdp_mean > 25000 | #<<
                                  health_mean < 1500 | #<<
                                  country %in% "Belgium"), #<<
                  mapping = aes(label = country))

```

---

# Label Outliers

.center[
```{r plot-outliers, ref.label = 'outliers', echo = FALSE, eval = TRUE}

```
]

---


```{r annotate, include = FALSE}
p <- ggplot(data = organdata,
            mapping = aes(x = roads,
                          y = donors))

p + 
  geom_point() +
  annotate(geom = "rect", 
           xmin = 125, xmax = 155, 
           ymin = 30, ymax = 35, 
           fill = "red", alpha = 0.2) + 
    annotate(geom = "text", 
           x = 91, 
           y = 33, 
           label = "A surprisingly high \n recovery rate", 
           hjust = 0) 

```


`r chunk_reveal(chunk_name = "annotate", title = "#Annotations", widths = c(50,50))`

---

class: center, middle, inverse

# Lab Time

### Pages 115 - 131 , 132 (bullet point 1) 



---

# Homework Assignment

**Task: ** Modify 3 previous plots with additional labels/annotations.  

**Due: ** March 3, 2021 

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42755_1&course_id=_35307_1)

**Notes**

- Use plots from labs or homework assignments

- Use your resources: Healy, Google, Student Community BUT cite where you get code from if you copy it directly  

- See Blackboard assignment for a template `.Rmd` file

---

# Midterm

**Task: ** Using R and an approved dataset, create a polished infographic that contains at least 3 charts.  

**Due: ** March 3, 2021

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42756_1&course_id=_35307_1)

- Create summary statistics and 3-5 polished visuals based on the dataset using **only R**. 

- The visuals created in R must include: titles, axis labels, data labels and legends (as needed), your name and the data source (as footnotes/captions)  

- Output the visuals to `.svg` and import them into Adobe Illustrator to create an infographic that includes a narrative, the visuals from R, and other graphics.  

- Turn in your R code as an `.Rmd` file, including comments about what you did and why, **and** the final infographic. 
- In class on **March 3**, be prepared to share information about your dataset, your visuals, what they show, and the choices you made in their design. You will share both your code and the final infographic.

???

In class on March 3, we will be sharing critiques of each others work. If you'd like to resubmit and make changes after class on March 3, you can do so and re-submit the assignment on Blackboard. 

Presentation to the class: share your .Rmd file and be prepared to explain your data (what is it about, what does it tell you, what messages did you find in it), your visualizations (how did you create them, why did you create them, what design choices you made); share your infographic (what design choices did you make and why, what story or message does the infographic show)
---

# Tasks to Complete

- Homework #3

- Midterm


