---
title: "GRD 610A Data Visualization II"
subtitle: "Summarizing Data"
author: "Jenn Schilling"
date: "March 17, 2021"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, bottom-note.css]
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

options(knitr.duplicate.label = "allow")

knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)

```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 80%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 10: Mapping Data in *The Truthful Art* (Cairo)  

- Lab on building histograms and box plots, data distributions, error and error bars  
(Section 4.6 and 5.2 of *Data Visualization* by Healy)  

- Homework Assignment #5


---

# Mapping Data

**Main Attributes of a Map:**

- Scale  
- Projection  
- Symbols 

.bottom-note[
(Cairo, 2016, p. 263)
]

???

Large-scale map (close-up of a  town) versus small-scale map (entire Earth)

Think about scale in relation to audience - how much do they know about the area and location?

Projection: "The process of making a globe, or a portion of it, fit into a flat picture." (Cairo, 2016, p. 265)
Some features will always be distorted (stretched or compressed) in the transformation


---

# Mapping Data

**Projection:**

- Developable surfaces   
- Standard lines 

**Distortions:**
- Shape  
- Area  
- Angles  
- Distance  
- Direction  

**Figure 10.4** shows and the text around it explains some differences in map projections

**Figure 10.6** provides a series of questions and a suggestion of map projection

.bottom-note[
(Cairo, 2016, pp. 265-267)
]

???

Developable surfaces - "the geometrical objects where the globe can be projected"

Standard lines - tangent to the surface; "scale is only accurate along those lines"; move further away from the standard lines -> more distortion

---

# Mapping Data

What were the key takeaways for you on mapping data?

???

Correctly sizing symbols is important

Viewers won't be able to estimate shape sizes very well

Symbol placement is important

Never trust software defaults (p. 280)

Take the time to set your own scale for color (p. 282-283)

Use choropleth maps in combination with each other; add other charts/tables to go with choropleth maps

Consider alternatives to area - circles or cartograms

--

Which maps did you find most engaging in this chapter? Why?

--

Which maps did you find least engaging in this chapter? Why?

--

What questions do you have?


---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "25%",
          warn_when = 60)
```

---

# Data: `midwest`

.center[
```{r midwest-data, echo = FALSE}
sample_n(midwest %>% select(state, county, area, percollege), size = 10) %>%
  knitr::kable(format = "html")
```
]

???

Some of the columns from the `midwest` dataset

percollege is percent college educated

---

# Histograms

.pull-left[
```{r histogram, include = FALSE}

p <- ggplot(data = midwest,
            mapping = aes(x = area))

p + 
  geom_histogram(bins = 10)

```

```{r decorate-histogram, echo = FALSE}

decorate("histogram", eval = FALSE) %>%
flair("geom_histogram") %>% 
knit_print.with_flair()

```
]

.pull-right[
```{r output-histogram, ref.label = 'histogram', echo = FALSE}
```
]

???

This plot provides the distribution of the area of counties in the midwest

---

# Density Plots

.pull-left[
```{r density, include = FALSE}

p <- ggplot(data = midwest,
            mapping = aes(x = area))

p + 
  geom_density()

```

```{r decorate-density, echo = FALSE}

decorate("density", eval = FALSE) %>%
flair("geom_density") %>% 
knit_print.with_flair()

```
]

.pull-right[
```{r output-density, ref.label = 'density', echo = FALSE}
```
]

???

`geom_density()` computes the kernel density estimate of the underlying distribution of the data

Exact same mapping as above; note the difference in the y-axis range

---

# Density Plots

.pull-left[
```{r density-state, include = FALSE}

p <- ggplot(data = midwest,
            mapping = aes(x = area,
                          fill = state,
                          color = state))

p + 
  geom_density(alpha = 0.3)

```

```{r decorate-density-state, echo = FALSE}

decorate("density-state", eval = FALSE) %>%
flair("geom_density") %>% 
knit_print.with_flair()

```
]

.pull-right[
```{r output-density-state, ref.label = 'density-state', echo = FALSE}
```
]

???

Can also fill density by state

---

# Data: `organdata`

.center[
```{r organ-data, echo = FALSE}
sample_n(organdata %>% select(country, year, donors), size = 10) %>%
  knitr::kable(format = "html")
```
]

???

Some of the columns from the `organdata` dataset


---

# Box Plots

.pull-left[
```{r organ-boxplot, include = FALSE}

p <- ggplot(data = organdata,
            mapping = 
              aes(x = reorder(country,
                              donors, 
                              na.rm = TRUE),
                  y = donors))


p + geom_boxplot() +
  labs(x = "") + 
  coord_flip()

```

```{r decorate-organ-boxplot, echo = FALSE}

decorate("organ-boxplot", eval = FALSE) %>%
flair("reorder") %>%
flair("geom_boxplot") %>% 
knit_print.with_flair()

```
]

.pull-right[
```{r output-organ-boxplot, ref.label = 'organ-boxplot', echo = FALSE}
```
]

---

# Summarize Data

```{r by-country}

by_country <- organdata %>%
  group_by(consent_law, country) %>%
  summarize(donors_mean = mean(donors, na.rm = TRUE),
            donors_sd = sd(donors, na.rm = TRUE))

```

.center[
```{r by-country-display, echo = FALSE}

head(by_country) %>%
  knitr::kable(format = "html")

```
]

---

# Point Range Plots

.pull-left[
```{r organ-pointrange, include = FALSE}

p <- ggplot(data = by_country,
            mapping = aes(x = reorder(country, 
                                      donors_mean),
                          y = donors_mean))

p + 
  geom_pointrange(
    mapping = 
      aes(ymin = donors_mean - donors_sd,
          ymax = donors_mean + donors_sd)) +
  labs(x = "",
       y = "Donor Procurement Rate") +
  coord_flip() 

```

```{r decorate-organ-pointrange, echo = FALSE}

decorate("organ-pointrange", eval = FALSE) %>%
flair("reorder") %>%
flair("geom_pointrange") %>% 
knit_print.with_flair()

```
]

.pull-right[
```{r output-organ-pointrange, ref.label = 'organ-pointrange', echo = FALSE}
```
]


???

Plot the dots (which represent the mean) with the range of standard deviation using `geom_pointrange()`

Need to use `coord_flip()` because `geom_pointrange()` uses `y`, `ymin`, and `ymax` and we want to show this on the x-axis



---

class: center, middle, inverse

# Lab Time

### Sections 4.6 and 5.2 


---

# Homework Assignment

**Task: ** Create 3 summary plots using data you have found  

**Due: ** March 24, 2021 

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42755_1&course_id=_35307_1)

**Notes**

- Use your own dataset (start looking for a dataset that you could use for the Final Project or go back to the dataset from the Midterm Project)

- Use your resources: Healy, Google, Student Community BUT cite where you get code from if you copy it directly  

- See Blackboard assignment for a template `.Rmd` file

---

# Tasks to Complete

- Reading (Cairo - Chapter 16: Exploring Data with Simple Charts)

- Homework #5

- Prepare for your Data Visualization of the Week




