---
title: "GRD 610A Data Visualization II"
subtitle: "Summarizing Data"
author: "Jenn Schilling"
date: "March 17, 2021"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, bottom-note.css]
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

options(knitr.duplicate.label = "allow")

knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)

```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 80%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 10: Mapping Data in *The Truthful Art* (Cairo)  

- Lab on building histograms and box plots, data distributions, error and error bars  
(Section 4.6 and 5.2 of *Data Visualization* by Healy)  

- Homework Assignment #5


---

# Maps

**"How do you know that you have chosen the right graphic form to represent your data?"**

.bottom-note[
(Cairo, 2016, pp. 121, 124-125)
]


---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "25%",
          warn_when = 60)
```

---

# New Dataset

**2016 U.S. Presidential Election Data**

.center[
```{r elections data, echo = FALSE}

election %>% 
  select(state, total_vote, r_points, pct_trump, party, census) %>%
  sample_n(10) %>%
  knitr::kable(format = "html")

```
]


---

class: center, middle, inverse

# Lab Time

### Sections 4.6 and 5.2 


---

# Homework Assignment

**Task: ** Create 3 summary plots using data you have found  

**Due: ** March 17, 2021 

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42755_1&course_id=_35307_1)

**Notes**

- Use your own dataset (start looking for a dataset that you could use for the Final Project)

- Use your resources: Healy, Google, Student Community BUT cite where you get code from if you copy it directly  

- See Blackboard assignment for a template `.Rmd` file

---

# Tasks to Complete

- Reading (Cairo - Chapter 16: Exploring Data with Simple Charts)

- Homework #5

- Prepare for your Data Visualization of the Week




