---
title: "GRD 610A Data Visualization II"
subtitle: "Effective Use of Color"
author: "Jenn Schilling"
date: "March 24, 2021"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, bottom-note.css]
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

options(knitr.duplicate.label = "allow")

knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)

```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 80%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 6: Exploring Data with Simple Charts in *The Truthful Art* (Cairo)  

- Lab on Color palettes for visualization. Heat Maps. Drawing attention effectively in a visualization. 
(??? of *Data Visualization* by Healy)  

- Homework Assignment #6


---

class: center, middle

### "The greatest value of a picture is when it forces us to notice what we never expect to see."
John W. Tukey, *Exploratory Data Analysis*

.bottom-note[
(Cairo, 2016, p. 151) 
]

???

Tukey created **exploratory data analysis** and says that "before you can even being testing your ideas against evidence, it's essential to get a good feel for what your data look like."

"the best way of doing so [getting a good feel for what you data look like] involves graphical displays, not just numerical summaries"

(Cairo, 2016, p. 151)

Cairo recommends finding data on a topic you're interested in and practicing with that

---
# Exploring Data with Simple Charts

"The process of visually exploring data can be summarized in a single sentence: **find patterns and trends lurking in the data and then observe the deviations from those patterns**."

.bottom-note[
(Cairo, 2016, p. 152) 
]


--

**Measures of Central Tendency**
- Mode  
- Median  
- Mean
- Weighted Mean

???

Mode: most frequent/common number
Median: middle number/number that divides dataset in half
Mean: average/sum divided be the number of numbers (sensitive to extreme values)

Example: Salaries from UNC Chapel Hill w/ Michael Jordan (pp. 154-155)

If Mean and Median are largely different, then that's a sign of skew

Weighted mean (p. 156 - 157)

--

**Measures of Spread**
- Range
- Shape/Histogram
- Plot

???

Spread gives us an idea how close the data is to the central measures
Range: the difference between the maximum and minimum values
Histogram: shows the frequency of each value or group of values; experiment with bin sizes
Look for skew

Making plots of the range and spread of data can help see trends

---

class: center, middle

### "Finding stories is sometimes a matter of repeatedly asking ourselves what would happen if we plot our data in a different way."

.bottom-note[
(Cairo, 2016, p. 160) 
]

???

The histogram/violin (Fig. 6.11), boxplots, pointrange (Fig. 6.9), and jitter plots (Fig. 6.10) we learned about in Week 9 are used in the figures in this section of the chapter.

---

class: center, middle

### "a visualization designer should never rely on a single statistic or single chart or map when doing exploratory work"

.bottom-note[
(Cairo, 2016, p. 163) 
]

---

# Exploring Data with Simple Charts

- What is your main takeaway from this chapter?

--

- Which visual representations make the most sense to you?

--

- Why is this important?

---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "25%",
          warn_when = 60)
```

---

# Data: `midwest`




---

class: center, middle, inverse

# Lab Time

### XXXX


---

# Homework Assignment

**Task: ** Create 3 visualizations that demonstrate effective color use  

**Due: ** March 31, 2021 

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42755_1&course_id=_35307_1)

**Notes**

- Use your own dataset (start looking for a dataset that you could use for the Final Project or go back to the dataset from the Midterm Project)

- Use your resources: Healy, Google, Student Community BUT cite where you get code from if you copy it directly  

- See Blackboard assignment for a template `.Rmd` file

---

# Tasks to Complete

- Reading (Cairo - CChapter 7: Visualizing Distributions)

- Homework #6

- Prepare for your Data Visualization of the Week




