---
title: "GRD 610A Data Visualization II"
subtitle: "Effective Use of Color"
author: "Jenn Schilling"
date: "March 24, 2021"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, bottom-note.css]
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

options(knitr.duplicate.label = "allow")

knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)
library(ggrepel)

```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Montserrat"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 70%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 6: Exploring Data with Simple Charts in *The Truthful Art* (Cairo)  

- Lab on color palettes for visualization, heat maps, and drawing attention effectively in a visualization. 

- Homework Assignment #6


---

class: center, middle

### "The greatest value of a picture is when it forces us to notice what we never expect to see."
John W. Tukey, *Exploratory Data Analysis*

.bottom-note[
(Cairo, 2016, p. 151) 
]

???

Tukey created **exploratory data analysis** and says that "before you can even being testing your ideas against evidence, it's essential to get a good feel for what your data look like."

"the best way of doing so [getting a good feel for what you data look like] involves graphical displays, not just numerical summaries"

(Cairo, 2016, p. 151)

Cairo recommends finding data on a topic you're interested in and practicing with that

---
# Exploring Data with Simple Charts

"The process of visually exploring data can be summarized in a single sentence: **find patterns and trends lurking in the data and then observe the deviations from those patterns**."

.bottom-note[
(Cairo, 2016, p. 152) 
]


--

**Measures of Central Tendency**
- Mode  
- Median  
- Mean
- Weighted Mean

???

Mode: most frequent/common number
Median: middle number/number that divides dataset in half
Mean: average/sum divided be the number of numbers (sensitive to extreme values)

Example: Salaries from UNC Chapel Hill w/ Michael Jordan (pp. 154-155)

If Mean and Median are largely different, then that's a sign of skew

Weighted mean (p. 156 - 157)

--

**Measures of Spread**
- Range
- Shape/Histogram
- Plot

???

Spread gives us an idea how close the data is to the central measures
Range: the difference between the maximum and minimum values
Histogram: shows the frequency of each value or group of values; experiment with bin sizes
Look for skew

Making plots of the range and spread of data can help see trends

---

class: center, middle

### "Finding stories is sometimes a matter of repeatedly asking ourselves what would happen if we plot our data in a different way."

.bottom-note[
(Cairo, 2016, p. 160) 
]

???

The histogram/violin (Fig. 6.11), boxplots, pointrange (Fig. 6.9), and jitter plots (Fig. 6.10) we learned about in Week 9 are used in the figures in this section of the chapter.

---

class: center, middle

### "a visualization designer should never rely on a single statistic or single chart or map when doing exploratory work"

.bottom-note[
(Cairo, 2016, p. 163) 
]

---

# Exploring Data with Simple Charts

- What is your main takeaway from this chapter?

--

- Which visual representations make the most sense to you?

--

- Why is this important?

--

- What questions do you have?

---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "25%",
          warn_when = 60)
```

---

# Data: `county_data`

.center[
```{r midwest-data, echo = FALSE}
sample_n(county_data %>% select(state, name, pop, black, flipped, partywinner16), size = 10) %>%
  knitr::kable(format = "html")
```
]

???

Lab today is about adding highlights and color/text effectively as well as changing theme elements and making heat maps

---
```{r plot-county-election, include = FALSE}

party_colors <- c("#2E74C0", "#CB454A")

ggplot(data = subset(county_data,
                           flipped == "No"),
             mapping = aes(x = pop,
                           y = black / 100)) +
  geom_point(alpha = 0.15,
             color = "gray50") +
  scale_x_log10(labels = scales::comma) +
  geom_point(data = subset(county_data,
                           flipped == "Yes"),
             mapping = aes(x = pop,
                           y = black / 100,
                           color = partywinner16)) +
  scale_color_manual(values = party_colors) +
  scale_y_continuous(labels = scales::percent) +
  labs(color = "County flipped to...",
       x = "County Population (log scale)",
       y = "Percent Black Population",
       title = "Flipped counties, 2016",
       caption = "Counties in gray did not flip.") +
  geom_text_repel(data = subset(county_data,
                                flipped == "Yes" & black > 25),
                  mapping = aes(x = pop,
                                y = black / 100,
                                label = state),
                  size = 2,
                  seed = 1234) +
  theme_minimal() +
  theme(legend.position = "top")

```

`r chunk_reveal(chunk_name = "plot-county-election", title = "#2016 Election Flipped Counties", widths = c(50,50))`

???
Initial plot - points for each county that did not flip  
x-axis is the population  
y-axis is the percent of the population that is Black  
Add a second layer of points for the counties that did flip  
color by party that won  
set y-axis and scales  
Label counties with high percentage of Black residents that flipped  
Add a minimal theme and put the legend at the top  



```{r make-plot, include = FALSE}
# Democrat Blue and Republican Red - create color palette
party_colors <- c("#2E74C0", "#CB454A")

# Initial plot - points for each county that did not flip
# x-axis is the population
# y-axis is the percent of the population that is Black
p0 <- ggplot(data = subset(county_data,
                           flipped == "No"),
             mapping = aes(x = pop,
                           y = black / 100))
p1 <- p0 +
  geom_point(alpha = 0.15,
             color = "gray50") +
  scale_x_log10(labels = scales::comma) # log-scale for population

# Add a second layer of points for the counties that did flip
# color by party that won
p2 <- p1 +
  geom_point(data = subset(county_data,
                           flipped == "Yes"),
             mapping = aes(x = pop,
                           y = black / 100,
                           color = partywinner16)) +
  scale_color_manual(values = party_colors)

# Set the y-axis scale and labels
p3 <- p2 +
  scale_y_continuous(labels = scales::percent) +
  labs(color = "County flipped to...",
       x = "County Population (log scale)",
       y = "Percent Black Population",
       title = "Flipped counties, 2016",
       caption = "Counties in gray did not flip.")

# Label counties with high percentage of Black residents that flipped
p4 <- p3 +
  geom_text_repel(data = subset(county_data,
                                flipped == "Yes" & black > 25),
                  mapping = aes(x = pop,
                                y = black / 100,
                                label = state),
                  size = 2)

```

---

# Changing Themes

.pull-left[
```{r theme-bw, eval = FALSE}

p4 + 
theme_bw() + #<<
theme(legend.position = "top")

```
]
.pull-right[
```{r output-theme-bw, ref.label = 'theme-bw', echo = FALSE}
```
]

???

Note: The book uses `theme_set()` which allows you to set the theme for all plots in a session. I'm going to use + `theme_xx()` because I don't want to set the theme for all the plots

---

# Changing Themes

.pull-left[
```{r theme-dark, eval = FALSE}

p4 + 
theme_dark() + #<<
theme(legend.position = "top")

```
]
.pull-right[
```{r output-theme-dark, ref.label = 'theme-dark', echo = FALSE}
```
]


---

# Changing Themes

.pull-left[
```{r theme-econ, eval = FALSE}

library(ggthemes) #<<

p4 + 
theme_economist() + #<<
theme(legend.position = "top")

```
]
.pull-right[
```{r output-theme-econ, ref.label = 'theme-econ', echo = FALSE}
```
]

---

# Changing Themes

.pull-left[
```{r theme-wsj, eval = FALSE}

library(ggthemes) #<<

p4 + 
theme_wsj() + #<<
theme(plot.title = element_text(size = rel(0.6)),
      legend.title = element_text(size = rel(0.35)),
      plot.caption = element_text(size = rel(0.35)),
      legend.position = "top")

```
]
.pull-right[
```{r output-theme-wsj, ref.label = 'theme-wsj', echo = FALSE}
```
]

???

Text sizes are quite large on this theme, so set them in the theme() function to be smaller

---

# Changing Themes

.pull-left[
```{r theme-custom, eval = FALSE}

p4 + 
theme(legend.position = "top",
      plot.title = element_text(size=rel(2),
                                lineheight=.5,
                                family="serif",
                                face="bold.italic",
                                color="orange"),
      axis.text.x = element_text(size=rel(1.1),
                                 family="mono",
                                 face="bold",
                                 color="purple"))

```
]
.pull-right[
```{r output-theme-custom, ref.label = 'theme-custom', echo = FALSE}
```
]

???

Make a custom theme  
Warnings - font family not found in Windows font database  
Needed to update the family names  

---

# Heat Map - Data

```{r hm-data, include = FALSE}

incidents <- read_csv("https://raw.githubusercontent.com/lgellis/MiscTutorial/master/ggmap/i2Sample.csv") 

# Prepare data - get month/year/day/etc. using lubridate package
incidents$ymd <-lubridate::mdy_hms(incidents$Event.Clearance.Date)
incidents$month <- lubridate::month(incidents$ymd, label = TRUE)
incidents$year <- lubridate::year(incidents$ymd)
incidents$wday <- lubridate::wday(incidents$ymd, label = TRUE)
incidents$hour <- lubridate::hour(incidents$ymd)

```

```{r hm-create-summary, include = FALSE}

# Create a summary table that has the number of incidents by hour and day
dayHour <- incidents %>%
  group_by(hour, wday) %>%
  summarise(N = n())

# Order by weekday in reverse
dayHour$wday <- factor(dayHour$wday, levels = rev(levels(dayHour$wday)))

# Remove NAs
dayHour <- dayHour %>% na.omit()

```

.center[
```{r heatmap-data, echo = FALSE}
dayHour[1:10,] %>%
  knitr::kable(format = "html")
```
]

---

```{r heatmap, include = FALSE}
ggplot(data = dayHour, 
       mapping = aes(x = hour, 
                     y = wday)) + 
  geom_tile(aes(fill = N), 
            colour = "white") + 
  scale_fill_gradient(low = "#d8e1cf",  
                      high = "#438484") +  
  theme_minimal() + 
  labs(title = "Heat Map of Seattle Incidents by Day of Week and Hour",
       x = "Incidents Per Hour", 
       y = "Day of Week",
       fill = "Total Incidents") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

```

`r chunk_reveal(chunk_name = "heatmap", title = "#Heat Map", widths = c(50,50))`


???

Source: https://www.littlemissdata.com/blog/heatmaps
Code adapted slightly from the source to match dpylr usage and things we've learned in class

---

class: center, middle, inverse

# Lab Time

### Sections 8.0 - 8.4
### <a href="https://rpubs.com/jennschilling/grd610a_w21_w10_notes" style="color:#cf7270">Heat Maps</a>

---

# Homework Assignment

**Task: ** Create 3 visualizations that demonstrate effective color use  

**Due: ** March 31, 2021 

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42755_1&course_id=_35307_1)

**Notes**

- Use your own dataset (start looking for a dataset that you could use for the Final Project or go back to the dataset from the Midterm Project)

- Use your resources: Healy, Google, Student Community BUT cite where you get code from if you copy it directly  

- See Blackboard assignment for a template `.Rmd` file

---

# Tasks to Complete

- Reading (Cairo - Chapter 7: Visualizing Distributions)

- Homework #6

- Prepare for your Data Visualization of the Week




