---
title: "GRD 610A Data Visualization II"
subtitle: "Data Manipulation"
author: "Jenn Schilling"
date: "February 17, 2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)
 
```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 80%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 3: The Truth Continuum of *The Truthful Art* (Cairo)  

- Lab on Summary statistics, grouped analysis, manipulating rows, columns and tables - Chapter 5 of *Data Visualization* (Healy)  

- Homework Assignment #2

---

# Discussion - The Truth Continuum

.pull-left[

#### Axiom: Any visualization is a model  
(Cairo, 2016, p. 69)

#### Coda: The more adequately a model fits whatever it stands for without being needlessley complex, and the easier it is for is intended audience to interpret correctly, the better it will be.  
(Cairo, 2016, p. 70)

]

--

.pull-right[

- How do you define a model?

{{content}}

]

--

- What does it mean for one model to be "better" than another?

{{content}}

--

- Do you agree or disagree with this axiom and coda? Why or why not?

{{content}}

--

- Why is it important to consider complexity, controls, and models in data visualization? 


???

"A model is an abstraction that describes, explains, or predicts something about the workings of nature" (p. 86)



---

# Discussion - The Truth Continuum


.pull-left[

#### "It's more complicated than that."

#### Good visualizations shouldn't over-simplify information. They need to clarify it. In many cases, clarifying a subject requires *increasing* the amount of information, not *reducing* it.

(Cairo, 2016, p. 78)

]

--

.pull-right[

- What does it mean to present nuance and context in data visualization? What are the implications of this on design?

{{content}}

]

--

- How does this relate to considering your audience when creating a data visualization?


???

It is important to consider the audience's prior knowledge, context, assumptions, and perspectives.


---

# Discussion - The Truth Continuum

### Mind Bugs

--

#### Patternicity

We look for and see patterns in everything, even when they are not there.

--

#### Storytelling

We like to generalize and see cause and effect, even when it's not there.

--

#### Confirmation

We look for information and interpret evidence as confirming our views, avoiding dissonance and ignoring alternate viewpoints.

--

.center[

#### What does this mean for data visualization design?

]

---

# Discussion - The Truth Continuum

- What do you think of Cairo's truth continuum? 

--

- How can we judge or evaluate the data visualizations we create on the truth continuum?

???

Judge based on rigor, effectiveness, precision, accuracy

--

- What should we aim for on the truth continuum? 

--

- What are some strategies that you will use to make truer data visualization?

???

Strategies: 
- use multiple figures
- consider the comparison (to what, to whom, to when, to where)
- take error into account (sampling error, etc.)
- increase depth
- increase breadth 
- look for other sources 

--

- Is there a conflict between simplifying and making true models/visualizations? How can a designer approach such a conflict?

---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "27%",
          warn_when = 60)
```


---
# A Reminder of The ggplot Process

1. Tell the `ggplot()` function what our data is. (`data = ... `)

2. Tell `ggplot()` what relationships we want to see.For convenience we will put the results of the first two steps in an object called `p`.  (`mapping = aes(...)`)

3. Tell ggplot how we want to see the relationships in our data.  (`geom_...`)

4. Layer on geoms as needed, by adding them to the `p` object one at a time.  

5. Use some additional functions to adjust scales, labels, tick marks, titles. 

From p. 60 of Healy, K. (2019). *Data visualization: A practical introduction*. Princeton ; Oxford: Princeton University Press.

---

# Chapter 5: Show the Right Numbers



---

# Grouping

.pull-left[
```{r group, fig.show = 'hide', include = FALSE}
p <- ggplot(data = gapminder,
            mapping = aes(x = year,
                          y = gdpPercap)) 

p + geom_line(mapping = aes(group = country)) 
  
```

```{r decorate-group, echo = FALSE}

  decorate("group", eval = FALSE) %>%
  flair("mapping = aes(group = country)") %>% 
  knit_print.with_flair()
  
  
```
]

.pull-right[
```{r plot-group, ref.label = 'group', echo = FALSE}
```
]





---

class: center, middle, inverse

# Lab Time


---

# Homework Assignment

**Task: ** Create 3 calculated fields and plot them.  

**Due: ** February 24, 2020 

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42755_1&course_id=_35307_1)

**Notes**

- You should explore a dataset other than the `gapminder` one we've been using in the labs (some ideas: `babynames`, `palmerpenguins`, a CSV file you found)  

- To use an R package dataset, remember to run `install.packages("package_name")` once in the console and add `library(package_name)` to the `setup` portion of the .Rmd file

- Use your resources: Healy, Google, Student Community BUT cite where you get code from if you copy it directly  

- See Blackboard assignment for a template .Rmd file

---

# Tasks to Complete

- Reading (Cairo - Chapter 4: Of Conjuectures and Uncertainty)  

- Homework #2 

- Prepare for your Data Visualization of the Week  


