---
title: "GRD 610A Data Visualization II"
subtitle: "Data Manipulation"
author: "Jenn Schilling"
date: "February 17, 2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)
 
```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 80%} # smaller code font size

```

# Today

- Data Visualization of the Week  

- Discussion of Chapter 3 of *The Truthful Art* (Cairo)  

- Lab on Summary statistics, grouped analysis, manipulating rows, columns and tables - Chapter 5 of *Data Visualization* (Healy)  

- Second Homework Assignment

---

# Discussion - The Truth Continuum



---

# Discussion




---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "27%",
          warn_when = 60)
```


---
# A Reminder of The ggplot Process

1. Tell the `ggplot()` function what our data is. (`data = ... `)

2. Tell `ggplot()` what relationships we want to see.For convenience we will put the results of the first two steps in an object called `p`.  (`mapping = aes(...)`)

3. Tell ggplot how we want to see the relationships in our data.  (`geom_...`)

4. Layer on geoms as needed, by adding them to the `p` object one at a time.  

5. Use some additional functions to adjust scales, labels, tick marks, titles. 

From p. 60 of Healy, K. (2019). *Data visualization: A practical introduction*. Princeton ; Oxford: Princeton University Press.

---

# Chapter 5: Show the Right Numbers



---

# Grouping

.pull-left[
```{r group, fig.show = 'hide', include = FALSE}
p <- ggplot(data = gapminder,
            mapping = aes(x = year,
                          y = gdpPercap)) 

p + geom_line(mapping = aes(group = country)) 
  
```

```{r decorate-group, echo = FALSE}

  decorate("group", eval = FALSE) %>%
  flair("mapping = aes(group = country)") %>% 
  knit_print.with_flair()
  
  
```
]

.pull-right[
```{r plot-group, ref.label = 'group', echo = FALSE}
```
]

---



---

```{r build-plot, include = FALSE}

p <- ggplot(data = gapminder,
            mapping = aes(x = year,
                          y = gdpPercap))

p + geom_line(color = "gray70",
              aes(group = country)) +
  geom_smooth(size = 1.1, 
              method = "loess", 
              se = FALSE) +
  scale_y_log10(labels = scales::dollar) +
  facet_wrap(~continent, ncol = 5) +
  labs(x = "Year",
       y = "GDP per capita",
       title = "GDP per capita on Five Continents")

```

`r chunk_reveal(chunk_name = "build-plot", title = "# Building a Plot", widths = c(50,50))`



---

class: center, middle, inverse

# Lab Time


---

# Homework Assignment

**Task: ** Create 3 calculated fields and plot them.  

**Due: ** February 24, 2020 

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42755_1&course_id=_35307_1)

**Notes**

- You should explore a dataset other than the `gapminder` one we've been using in the labs (some ideas: `babynames`, `palmerpenguins`, a CSV file you found)  

- To use an R package dataset, remember to run `install.packages("package_name")` once in the console and add `library(package_name)` to the `setup` portion of the .Rmd file

- Use your resources: Healy, Google, Student Community BUT cite where you get code from if you copy it directly  

- See Blackboard assignment for a template .Rmd file

---

# Tasks to Complete

- Reading (Cairo - Chapter 4)  

- Homework #2 

- Prepare for your Data Visualization of the Week  


