---
title: "Data Visualization Notes"
author: "Jenn Schilling"
date: "2021-01-27"
output:
 html_document:
  theme: "flatly" # Theme Gallery: https://www.datadreaming.org/post/r-markdown-theme-gallery/
  toc: TRUE
  toc_float: TRUE
  code_download: TRUE
---
   
## Data Visualization Notes

These are my notes for **GRD 610A: Data Visualization II** in Winter 2021 at the College for Creative Studies. These notes are for my work in the book *Data Visualization* by Kieran Healy (Princeton University Press, 2019). 

```{r setup, include=FALSE}

## By defult, show code for all chunks in the knitted document,
## as well as the output. To override for a particular chunk
## use echo = FALSE in its options.
knitr::opts_chunk$set(echo = TRUE) 

## Set the default size of figures
knitr::opts_chunk$set(fig.width=8, fig.height=5)  

## Load the libraries we will be using
library(gapminder)
library(here)
library(socviz)
library(tidyverse)

## Libraries: Install once per machine, load once per R session

## Week 2 / Chapter 2 Notes published here: https://rpubs.com/jennschilling/dgr610a_w21_w2_notes

```


## Get Started

### Everything has a name
Objects in R are created and referred to by their names. Certain names are not allowed because they are reserved words such as ``TRUE``, ``if``, ``mean()``, and ``NA``. Names also cannot start with a number or contain spaces. There are different naming conventions.  

**Snake Case**  
``my_data``  
``this_is_snake_case``  

**Camel Case**  
``myData``  
``thisIsCamelCase``

**Pascal Case**  
``MyData``  
``ThisIsPascalCase``

Pick one naming convention and stick with it. Be consistent; don't switch between conventions. I recommend snake case.

```{r naming, error=TRUE}

# This is a comment (it starts with #)

my_data <- c(1, 2, 3, 4) # Assign using <- ; use ALT + - or OPTION + -

My_Data  
# Cannot be found because we called it my_data (lowercase)

# Now we can see it
my_data 

```

### Everything is an object; using functions
Think of functions like a recipe. The arguments of the function are the ingredients and what happens within the function is the sequence of cooking steps. 
```{r objects-functions}

c(1, 2, 3, 1, 3, 5, 25) # c() is the combine function, it puts things together into a vector/list

my_numbers <- c(1, 2, 3, 1, 3, 5, 25)
your_numbers <- c(5, 31, 71, 1, 3, 21, 6)

my_numbers

mean(x = my_numbers)
mean(my_numbers) # you don't have to specify the argument names, but order matters if you do not specify

mean(x = your_numbers)

my_summary <- summary(my_numbers)

my_summary

table(my_numbers)

sd(my_numbers)

my_numbers * 5

my_numbers + 1

my_numbers + my_numbers # How is this different than the last line?

# If you're not sure what an object is, ask for its class or type

class(my_numbers)

class(my_summary)

class(summary)

my_new_vector <- c(my_numbers, "Apple") # What happens if we combine a word with numbers?

my_new_vector

class(my_new_vector)

# Let's look at a new dataset

titanic

class(titanic) 

# Titanic is a data frame, which is like a table
# The $ operator can be used to access a column of a data frame by name

titanic$percent

# Tibbles are slightly different than data frames. They are also data tables, but they provide more information.

titanic_tb <- as_tibble(titanic)

titanic_tb # How is does this compare to titanic above?

# To see inside an object, ask for its structure

str(my_numbers)

str(my_summary)

```
Programming in R can be challenging and it takes time to get used to. Be patient and take a break if you get stuck. Make sure parentheses are opened and closed. Complete your commands (look out for the + in the console). Take your time and lookout for typos. 

### Get Data into R
In this section, we will get data from a URL and make a quick figure.

```{r get-data}

# Data source
url <- "https://cdn.rawgit.com/kjhealy/viz-organdata/master/organdonation.csv"

# Read the CSV from the URL
organs <- read_csv(file = url)

# Take a quick look at the data
glimpse(organs)

# View(organs) # Run in RStudio

```

```{r make-figure}

# Another way to view data 
gapminder

# Make a plot object
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap, 
                          y = lifeExp))

# Create a scatterplot
p + geom_point()

```


## Make a Plot

```{r}

```


## Show the Right Numbers

```{r}

```


## Graph Tables, Make Labels, Add Notes

```{r}

```

## Work with Models

```{r}

```

## Draw Maps

```{r}

```


## Refine your Plots

```{r}

```


