---
title: "GRD 610A Data Visualization II"
subtitle: "Midterm"
author: "Jenn Schilling"
date: "March 3, 2021"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, bottom-note.css]
    nature:
      ratio: '16:9'
      img_dir: img
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)

options(knitr.duplicate.label = "allow")

knitr::opts_chunk$set(fig.retina = 3, warning = FALSE, message = FALSE)

library(gapminder)
library(here)
library(socviz)
library(tidyverse)
library(countdown)
library(xaringanthemer)
library(flair)
library(flipbookr)
library(DT)
 
```

```{r xaringan-themer, include=FALSE, warning=FALSE}

style_duo_accent(
  primary_color = "#1c5253",
  secondary_color = "#c8c8c8",
  link_color = choose_dark_or_light("#FFFFFF", "#cf7270"),
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono"),
  inverse_background_color = "#1c5253",
  inverse_header_color = "#FFFFFF"
)

```

```{css change-css-options, eval = TRUE, echo = FALSE}

.remark-code{line-height: 1.5; font-size: 80%} # smaller code font size

```

# Today

- Midterm 

---

# Midterm

**Task: ** Using R and an approved dataset, create a polished infographic that contains at least 3 charts.  

**Due: ** March 3, 2021 **(TODAY)**

[Rubric](https://bb.collegeforcreativestudies.edu/webapps/rubric/do/course/manageRubrics?dispatch=view&context=course&rubricId=_42756_1&course_id=_35307_1)

- Create summary statistics and 3-5 polished visuals based on the dataset using **only R**. 

- The visuals created in R must include: titles, axis labels, data labels and legends (as needed), your name and the data source (as footnotes/captions)  

- Output the visuals to `.svg` and import them into Adobe Illustrator to create an infographic that includes a narrative, the visuals from R, and other graphics.  

- Turn in your R code as an `.Rmd` file, including comments about what you did and why, **and** the final infographic. 
- In class on **March 3**, be prepared to share information about your dataset, your visuals, what they show, and the choices you made in their design. You will share both your code and the final infographic.

???

In class on March 3, we will be sharing critiques of each others work. If you'd like to resubmit and make changes after class on March 3, you can do so and re-submit the assignment on Blackboard. 

Presentation to the class: share your .Rmd file and be prepared to explain your data (what is it about, what does it tell you, what messages did you find in it), your visualizations (how did you create them, why did you create them, what design choices you made); share your infographic (what design choices did you make and why, what story or message does the infographic show). 

---

class: center, middle

# 15 Minute Break

```{r break-countdown, echo = FALSE}
countdown(minutes = 15, seconds = 0,
          right = "37%",
          bottom = "25%",
          warn_when = 60)
```



---

# Tasks to Complete

- Reading (Cairo - Chapter 5: Basic principles of visualization)

- Prepare for your Data Visualization of the Week

- **[Optional]** Make changes to your midterm and resubmit

